---
title: "Build GRASS GIS database"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=2cm
output: 
  NinaR::jensAnalysis:
    highlight: tango
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
# This is optional
# I choose the 'styler' package for tidying the code to preserve indentations
# I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
# Set tidy = True to get the knitr default
# I want all figures as png and pdf in high quality in a subfolder called figure
require(knitr)
require(NinaR)

knitr::opts_chunk$set(
  eval = FALSE, # do not run anything
  echo = TRUE,
  tidy = "styler",
  dev = c("png", "pdf"),
  dpi = 600,
  fig.path = "figure/"
)

options(
  xtable.comment = F,
  xtable.include.rownames = F,
  nina.logo.y.pos = 0.15
)
palette(ninaPalette())
```

```{r load_packages, message = F, warning = F}
# Load packages
require(dplyr)
require(purrr)
require(readr)
require(NinaR)

require(rgrass7)
require(terra)
require(rgdal)
require(sf)

require(oneimpact)

require(RPostgres)
require(rpostgis)

source("functions.R")
```

# Connect to GRASS GIS and PostGIS

We start by connecting the R session to GRASS GIS, in my local user mapset (`u_bernardo.brandao`).

```{r connectGRASS}
# Connect to GRASS
NinaR::grassConnect(location = "ETRS_33N", mapset = "user")

# example of command
#execGRASS("g.mapset", parameters = list(), flags = )
```

We also connect to the `gisdata` PostGIS database.

```{r connectpostgis}
source("~/.pgpass")

postgreSQLConnect(
  host = "gisdata-db.nina.no",
  dbname = "gisdata",
  username = pg_username,
  password = pg_password
)

# check
con
pgPostGIS(con)

# make string to link to PostGIS data in data importing to GRASS
str_con <- paste0("PG:host=gisdata-db.nina.no user=", pg_username, " password=", pg_password, " dbname=gisdata")

# remove stored username and password
rm(pg_username, pg_password)
```

# Get read access to the existing mapsets

```{r}
# list all mapsets
sep <- " "
all_mapsets <- execGRASS("g.mapsets", flags = c("l"), intern = T) %>% 
  strsplit(sep) %>% 
  first()

# list mapsets for "rein"
mapset_patt_rein <- c("Rein", "rein") %>% 
  paste(collapse = "|")
mapsets_rein <- all_mapsets %>% 
  grep(pattern = mapset_patt_rein, value = T)

# list mapsets for "Prodchange"
mapset_patt_pc <- c("change") %>% 
  paste(collapse = "|")
mapsets_pc <- all_mapsets %>% 
  grep(pattern = mapset_patt_pc, value = T) %>% 
  grep(pattern = paste(mapsets_rein,collapse = "|"), invert = T, value = T)

# list some more mapsets we know we'll use
land_use_mapsets <- c("p_RenRein_norut", "g_LandCover_Norway_NORUT_SAM_TT", 
                      "p_prodchange_envpolyTT")
landscape_mapsets <- c("g_Elevation_Fenoscandia", "g_LandCover_Fenoscndia_PHENOLOGY_SAM_TT", 
                       "g_Elevation_Fenoscandia_TPI")
climate_mapsets <- c("g_BiogeographicalRegions_Norway_PCA_klima", "u_bram.van.moorter",
                     "g_EnergyResources_Fenoscandia", "u_torkildtveraa")
infrastructure_mapsets <- c("p_prodchange_envpointsTT", "p_prodchange_roadsTT", "p_RenRein_trails2", 
                            "p_prodchange_trailsTT")

all_relevant_mapsets <- c(mapsets_rein, mapsets_pc, 
                          land_use_mapsets, landscape_mapsets, 
                          climate_mapsets, infrastructure_mapsets) %>% 
  unique()

# access to those mapsets
execGRASS("g.mapsets", parameters = list(mapset = all_relevant_mapsets))
```

# Mapset organization

We organized the environmental layers in six main mapsets, to store and process maps related to
different themes:

- `sam_climate_phenology`: maps related to climate (e.g. temperature, precipitation, snow, 
bioclim variables or Norwegian PCAs) and phenology (e.g. NDVI)
- `sam_landscape`: maps related to land use and land cover (e.g. land cover, forestry, deforestation)
- `sam_species` (should we call it `animals` or `biotic interactions`?): maps related to the influence or
interaction with other species (e.g. domesticated species, predators, competitors)
- `sam_industry`: maps related to industrial activity (e.g. wind power development, mining)
- `sam_transport_urban`: maps related to transport (e.g. roads, railways) and 
urban infrastructure (e.g. urban areas)
- `sam_tourism`: maps related to touristic activity (e.g. private cottages, public cabins, trails)

Besides, there is another one for ancillary files:

- `sam_tools`: maps not related to environmental layers (e.g. grid of 100m across Norway, management areas, 
herding districts)

## Create mapsets

```{r create_mapsets}
# Create mapsets
mapsets <- c("climate_phenology", "landscape", "species", 
             "industry", "transport_urban", "tourism", "tools")
mapsets <- paste0("p_sam_", mapsets)

for(ms in mapsets) {
  execGRASS("g.mapset", parameters = list(mapset = ms), flags = "c")
}
```

# Ancillary data

## Limits (national, Sami herding districts, management areas)

```{r}
# go into mapset
ms <- "p_sam_tools"
execGRASS("g.mapset", mapset = ms)

crs_no <- 25833 #EPSG:25833 (ETRS89/UTM zone 33N)

# national limits

# norway
gisdata_schema <- "sam_tools"
layer_name <- "norway_limits"
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")
# import
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

# sweden
gisdata_schema <- "sam_tools"
layer_name <- "sweden_limits"
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")
# import
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

# Sami herding districts

# norway
gisdata_schema <- "sam_trein_ancillary"
layer_name <- "herding_districts_no_2012"
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")
# import
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

# sweden
gisdata_schema <- "sam_trein_se_ancillary"
layer_name <- "herding_districts_se"
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")
# import
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

gisdata_schema <- "sam_trein_se_ancillary"
layer_name <- "herding_districts_se_union"
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")
# import
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

# Wild reindeer management areas Norway

# official
gisdata_schema <- "sam_wrein_ancillary"
layer_name <- "reindeer_areas_official"

# read to R and reproject - it is currently in SWEREF99
reindeer_areas_official <- sf::st_read(con, Id(schema = gisdata_schema, 
                                               table = layer_name)) %>% 
  sf::st_transform(crs = crs_no) %>% 
  terra::vect()

# import to GRASS
rgrass7::write_VECT(reindeer_areas_official, "reindeer_areas_official_no")

# used in reindeer projects (renrein, prodchange)
gisdata_schema <- "sam_wrein_ancillary"
layer_name <- "reindeer_areas"

# read to R and reproject - it is currently in UTM 32N
reindeer_areas <- sf::st_read(con, Id(schema = gisdata_schema, 
                                      table = layer_name)) %>% 
  sf::st_transform(crs = crs_no) %>%
  terra::vect()

# import to GRASS
rgrass7::write_VECT(reindeer_areas, "reindeer_areas_no")
```

## Master grid for Norway

```{r master_grid}
# go into mapset
ms <- "p_sam_tools"
execGRASS("g.mapset", mapset = ms)

# root files
grid_file_1km = "/data/R/GeoSpatialData/Population_demography/Norway_SSB/Processed/master_grid_1km.tif"
grid_file_100m = "/data/R/GeoSpatialData/Population_demography/Norway_SSB/Processed/master_grid_100m.tif"

# write files to GRASS
rgrass7::execGRASS("r.in.gdal", input = grid_file_1km, output = "master_grid_1km_norway")
rgrass7::execGRASS("r.in.gdal", input = grid_file_100m, output = "master_grid_100m_norway")
```

# Landscape data

## Data from Norway

Here we'll copy the datasets from Norway that have already been in use in the previous projects (e.g. Renewable Reindeer, Prodchange).

### Land use - NORUT

```{r copy_landscape_NORUT}
# go into mapset
ms <- "p_sam_landscape"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "landscape"

#---
# land use
ms_from <- "p_RenRein_norut"
maps <- execGRASS("g.list", type = "raster", mapset = ms_from, intern = T)

# copy
for(i in maps) {
  map_in <- paste0(i, "@", ms_from)
  map_out <- i
  execGRASS("g.region", raster = map_in)
  execGRASS("g.copy", raster=paste0(map_in, ",", map_out), flags = c("overwrite"))
}

# document
md_file <- list.files("../data/", pattern = "spatialdb_metadata_oneimpact", full.names = T)
md <- readr::read_csv(md_file)

# maps used or not used earlier
used <- grep("100", maps, value = T)
not_used <- setdiff(maps, used)
used_names <- c("agricultural lands", "glacier", "grasses", "heather in lowland", "heather in ridges",
                "heathlands", "forest with lichens", "lichens", "meadows", "mires", "ridges", "forest",
                "snow", "snowbed")
used_names <- paste0("Land cover: ", c(used_names))
not_used_names <- used_names

# used maps
md_updated <- update_metadata(md, 
                              maps = used, 
                              type_of_info = type_of_info,
                              mapset_from = ms_from, 
                              new_mapset = ms, 
                              variables = used_names, 
                              institution = NA,
                              description = NA, 
                              unit = NA, 
                              type_data = "raster",
                              original_range_values = "0, 1", 
                              year_data = NA,
                              original_pixel_res = 30, #???
                              final_pixel_res = 30, 
                              extent = "Norway",
                              primary_derived = "Primary",
                              derived_form = NA,
                              website = NA, 
                              source = NA, 
                              obtained_through = "NINA", 
                              observations = NA)
md_updated

# maps unused earlier
md_updated <- update_metadata(md_updated, 
                              maps = not_used, 
                              type_of_info = type_of_info,
                              mapset_from = ms_from, 
                              new_mapset = ms, 
                              variables = not_used_names, 
                              institution = NA,
                              description = NA, 
                              unit = NA, 
                              type_data = "raster",
                              original_range_values = "0, 1", 
                              year_data = NA,
                              original_pixel_res = 30, #??
                              final_pixel_res = 100, 
                              extent = "Norway",
                              primary_derived = "Resampled",
                              derived_form = "NORUT land use classes",
                              website = NA, 
                              source = NA, 
                              obtained_through = "NINA", 
                              observations = NA)
md_updated

# write updated metadata on the disk
readr::write_csv(md_updated, file = paste0("../data/spatialdb_metadata_oneimpact_",
                                           gsub("-", "", lubridate::today()),
                                           ".csv"))
# put the old metadata file in archive
# system(paste("mv", md_file, "../data/old_metadata"))

#---
# land use density
# This data is present in the mapset "g_LandCover_Norway_NORUT_SAM_TT", 
# but it should be recalculated
# with the data from Sweden
# But we need to compare the land use maps first
```

### Elevation for Fennoscandia

```{r copy_landscape_elevation}
# go into mapset
ms <- "p_sam_landscape"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "landscape"

#---
# elevation, slope, aspect???
```

### TPI for Fennoscandia

```{r copy_landscape_TPI}
# go into mapset
ms <- "p_sam_landscape"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "landscape"

#---
# TPI - 50m
ms_from <- "g_Elevation_Fenoscandia_TPI"
maps <- execGRASS("g.list", type = "raster", mapset = ms_from, intern = T)

# copy
for(i in maps) {
  map_in <- paste0(i, "@", ms_from)
  map_out <- i
  execGRASS("g.region", raster = map_in)
  execGRASS("g.copy", raster=paste0(map_in, ",", map_out), flags = c("overwrite"))
}

# document
md_file <- list.files("../data/", pattern = "spatialdb_metadata_oneimpact", full.names = T)
md <- readr::read_csv(md_file)

maps
scales <- purrr::map_chr(strsplit(maps, "_"), ~.[3])
map_names <- paste0("Terrain Position Index: ", scales, "m")

# get min and max from GRASS maps
min_max <- get_univar(maps, vars = c("min", "max"), all_together = F)
min_max <- apply(min_max, 1, function(x) paste0("[", paste(round(x, 2), collapse = ", "), "]"))

md_updated <- update_metadata(md, 
                              maps = maps, 
                              type_of_info = type_of_info,
                              mapset_from = ms_from, 
                              new_mapset = ms, 
                              variables = map_names, 
                              institution = NA,
                              description = NA, 
                              unit = NA, 
                              type_data = "raster",
                              original_range_values = min_max, 
                              year_data = NA,
                              original_pixel_res = NA, #??
                              final_pixel_res = 50, 
                              extent = "Fenoscandia",
                              primary_derived = "Derived",
                              derived_form = "???",
                              website = NA, 
                              source = NA, 
                              obtained_through = "NINA", 
                              observations = NA)
md_updated

# write updated metadata on the disk
readr::write_csv(md_updated, file = paste0("../data/spatialdb_metadata_oneimpact_",
                                           gsub("-", "", lubridate::today()),
                                           ".csv"))
# put the old metadata file in archive
# system(paste("mv", md_file, "../data/old_metadata"))
```

### Digestible biomass

```{r copy_landscape_digbiomass}
# go into mapset
ms <- "p_sam_landscape"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "landscape"

#---
# Digestible biomass
ms_from <- "u_bram.van.moorter"
maps <- execGRASS("g.list", type = "raster", mapset = ms_from, intern = T) %>% 
  grep(pattern = "digest", value = T)

# copy
for(i in maps) {
  map_in <- paste0(i, "@", ms_from)
  map_out <- i
  execGRASS("g.region", raster = map_in)
  execGRASS("g.copy", raster=paste0(map_in, ",", map_out), flags = c("overwrite"))
}

# document
md_file <- list.files("../data/", pattern = "spatialdb_metadata_oneimpact", full.names = T)
md <- readr::read_csv(md_file)

maps
map_names <- c("Digestible biomass in summer", "Digestible biomass in winter")

# get min and max from GRASS maps
min_max <- get_univar(maps, vars = c("min", "max"), all_together = F)
min_max <- apply(min_max, 1, function(x) paste0("[", paste(round(x, 1), collapse = ", "), "]"))

md_updated <- update_metadata(md, 
                              maps = maps, 
                              type_of_info = type_of_info,
                              mapset_from = ms_from, 
                              new_mapset = ms, 
                              variables = map_names, 
                              institution = NA,
                              description = NA, 
                              unit = NA, 
                              type_data = "raster",
                              original_range_values = min_max, 
                              year_data = NA,
                              original_pixel_res = NA, #??
                              final_pixel_res = 100, 
                              extent = "Norway",
                              primary_derived = "Derived",
                              derived_form = "???",
                              website = NA, 
                              source = NA, 
                              obtained_through = "NINA", 
                              observations = NA)
md_updated

# write updated metadata on the disk
readr::write_csv(md_updated, file = paste0("../data/spatialdb_metadata_oneimpact_",
                                           gsub("-", "", lubridate::today()),
                                           ".csv"))
# put the old metadata file in archive
# system(paste("mv", md_file, "../data/old_metadata"))
```

## Data from Sweden

Here we load the landscape datasets from Sweden into GRASS GIS.

### Land use NMD, SMD, and Lichen model

```{r landscape_sweden}
# go into mapset
ms <- "p_sam_landscape"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "landscape"

#---
# folder
landscape_dir = paste0(sw_dir, "03_raster/sam_landscape/")

# load vegetation data - NMD
execGRASS("r.import", input = paste0(landscape_dir, "landcover_nmd_ungeneralized/nmd2018bas_ogeneraliserad_v1_0.tif"), 
    output = "landcover_sw_nmd1_ungeneralized_10m_2018")

# load auxiliary vegetation data - SMD
execGRASS("r.import", input = paste0(landscape_dir, "landcover_smd_generalized_2013/mosaic/smdb99.tif"), 
    output = "landcover_sw_smd_25m_2004")

# elevation, aspect, slope, tpi, 10m resampled from 2m - should I add?

# load general lichen map for Sweden - Sven
execGRASS("r.import", input = paste0(landscape_dir, "lichen_sven/lav_model_south_no_roads_masked.tif"),
    output = "lichen_model_sw", flags = "overwrite")

# update metadata

# read metadata from Swedish metadata
md_sw <- readr::read_csv("../data/spatialdb_metadata_sweden_20211101.csv")[1:19]
# read new metadata file
md_file <- list.files("../data/", pattern = "spatialdb_metadata_oneimpact", full.names = T)
md <- readr::read_csv(md_file)

# pattern to search in the Swedish db
patt <- c("ungeneralized", "smd_25m", "lichen_Sweden")

# find which lines correspond to these maps in the original sweden db
ind <- purrr::map_int(patt, ~ grep(., md_sw$`Layer name`))

# final map names  
maps <- c("landcover_sw_nmd1_ungeneralized_10m_2018", "landcover_sw_smd_25m_2004",
          "lichen_model_sw")

md_updated <- update_metadata_existing(md, maps = maps, indexes = ind, existing_metadata = md_sw)
md_updated

# write updated metadata on the disk
readr::write_csv(md_updated, file = paste0("../data/spatialdb_metadata_oneimpact_",
                                           gsub("-", "", lubridate::today()),
                                           ".csv"))
# put the old metadata file in archive
# system(paste("mv", md_file, "../data/old_metadata"))
```

We still need to process some data - doing operations like resampling for a single same resolution, mosaicing maps, calculating densities etc..



# Load climate and phenology data

Now we import climate and phenology related data.

## Data from Norway

Here we'll copy the datasets from Norway that have already been in use in the previous projects (e.g. Renewable Reindeer, Prodchange).

### Climate PCAs from Norway

```{r climate_pcas}
# go into mapset
ms <- "p_sam_climate_phenology"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "climate_phenology"

#---
# PCAs 1 and 2
ms_from1 <- "g_BiogeographicalRegions_Norway_PCA_klima"
maps1 <- execGRASS("g.list", type = "raster", mapset = ms_from1, intern = T) 

# PCAs 3 and 4
ms_from2 <- "u_bram.van.moorter"
maps2 <- execGRASS("g.list", type = "raster", mapset = ms_from2, intern = T) %>% 
  grep(pattern = "klima", value = T)

ms_from <- c(rep(ms_from1, 2), rep(ms_from2, 2))
maps <- c(maps1, maps2)

# copy
for(i in 1:length(maps)) {
  map_in <- paste0(maps[i], "@", ms_from[i])
  map_out <- maps[i]
  execGRASS("g.region", raster = map_in)
  execGRASS("g.copy", raster=paste0(map_in, ",", map_out), flags = c("overwrite"))
}

# document
md_file <- list.files("../data/", pattern = "spatialdb_metadata_oneimpact", full.names = T)
md <- readr::read_csv(md_file)

maps
map_names <- paste0("Norway climatic PCA", 1:4)

# get min and max from GRASS maps
min_max <- get_univar(maps, vars = c("min", "max"), all_together = F)
min_max <- apply(min_max, 1, function(x) paste0("[", paste(round(x, 4), collapse = ", "), "]"))

# description
desc <- paste("PCA axes from climatic variables for Norway, from 
Bakkestuen, V., Erikstad, L., and Halvorsen, R. (2008).
Step-less models for regional environmental variation in Norway. 
Journal of biogeography, 35(10):1906–1922.")
derived_fr <- paste("Fifty-four climatic, topographical, hydrological and geological. 
See Bakkestuen et al. (2008) for details")

md_updated <- update_metadata(md, 
                              maps = maps, 
                              type_of_info = type_of_info,
                              mapset_from = ms_from, 
                              new_mapset = ms, 
                              variables = map_names, 
                              institution = NA,
                              description = desc, 
                              unit = NA, 
                              type_data = "raster",
                              original_range_values = min_max, 
                              year_data = NA,
                              original_pixel_res = 1000, #??
                              final_pixel_res = 100, 
                              extent = "Norway",
                              primary_derived = "Derived",
                              derived_form = derived_fr,
                              website = NA, 
                              source = NA, 
                              obtained_through = "NINA", 
                              observations = NA)
md_updated

# write updated metadata on the disk
readr::write_csv(md_updated, file = paste0("../data/spatialdb_metadata_oneimpact_",
                                           gsub("-", "", lubridate::today()),
                                           ".csv"))
# put the old metadata file in archive
# system(paste("mv", md_file, "../data/old_metadata"))
```

### Phenology variables from Fennoscandia

The phenology variables seem to be available for the whole Fennoscandia (mapsets g_LandCover_Fenoscandia_PHENELOGY*).  
However, the variables currently available were created only for the extent of Norway, and I cannot find the files, 
mapsets (e.g. g_LandCover_Fenoscandia_PHENOLOGY_average does not exist in the server), or scripts used to create
them. Maybe we could re-calculate all that for the whole Fennoscandia?

Check that with Knut and Torkild.

## Data from Sweden

Here we load the landscape datasets from Sweden into GRASS GIS.








# Load species data

Now we import species/biotic related data.

## Data from Norway

Here we'll copy the datasets from Norway that have already been in use in the previous projects (e.g. Renewable Reindeer, Prodchange).

## Data from Sweden

### Predators?

```{r}
# go into mapset
execGRASS("g.mapset", parameters = list(mapset = "p_sam_species"))

# folder with raw data
species_dir <- paste0(sw_dir, "03_raster/p_sam_species/")

# list maps in folder
files <- list.files(species_dir, pattern = "tif$", full.names = T)

test <- terra::rast(files[1])
plot(test)

# load predators data into GRASS
for(i in file) {
  
  # prepare input name
  name <- strsplit(i, "/", fixed = T)[[1]] %>% 
    dplyr::last() %>% 
    gsub(pattern = ".tif", replacement = "")
  # execGRASS("r.in.gdal", parameters = list(input = i, output = name), flags = c("overwrite"))
  
  execGRASS("r.import", parameters = list(input = i, output = name), flags = c("overwrite"))
}
    
# document

```

# Industrial infrastructure

## Industrial areas and buildings

### Norway

```{r industrial_areas_no}
# go into mapset
ms <- "p_sam_industry"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "industry"

#---
# import temporary layer with industrial areas

# schema
gisdata_schema <- "sam_tmp"
# layer to use
layer_name <- "tmp_industrial_areas_no"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# import railways
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

#---
# import temporary layer with industrial buildings (points)

# schema
gisdata_schema <- "sam_tmp"
# layer to use
layer_name <- "tmp_industrial_buildings_no"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# import railways
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

#-----------------
# rasterize

# rasterize industrial areas with value 0.5
# region
rgrass7::execGRASS("g.region", raster = "master_grid_100m_norway@p_sam_tools", 
                   flags = "p")
# mask
rgrass7::execGRASS("r.mask", vector = "norway_limits@p_sam_tools")
# rasterize
rgrass7::execGRASS("v.to.rast", input = "tmp_industrial_areas_no", type = "area",
                   output = "tmp_industrial_areas_x05_no", use = "val", value = 0.5)

# rasterize houses counting the number of features per pixel
rgrass7::execGRASS("g.region", raster = "master_grid_100m_norway@p_sam_tools", 
                   flags = "p")
# rgrass7::execGRASS("r.mask", flags = "r")
oneimpact::grass_v2rast_count(x = "tmp_industrial_buildings_no", 
                              output = "tmp_industrial_buildings_count_no", 
                              set_region = FALSE)

#-----------------
# patch results

# region
rgrass7::execGRASS("g.region", raster = "master_grid_100m_norway@p_sam_tools", 
                   flags = "p")
# mask
rgrass7::execGRASS("r.mask", vector = "norway_limits@p_sam_tools")
# patch
rgrass7::execGRASS("r.patch", input = "tmp_industrial_areas_x05_no,tmp_industrial_buildings_count_no",
                   output = "industrial_buildings_count_no")
# export for checking
# dir <- "../data/sam_env/industry/industrial/processed/"
# rgrass7::execGRASS("r.out.gdal", input = "industrial_buildings_count_no", 
#                    output = paste0(dir, "industrial_buildings_count_no.tif"),
#                    createopt = "COMPRESS=DEFLATE")
# remove mask
rgrass7::execGRASS("r.mask", flags = "r")
```

### Sweden



# Transport and urban infrastructure

## Data from Sweden

Here we load the transport and urban infrastructure datasets from Sweden into GRASS GIS.

## Roads

### Norway

to do after checking summer/winter, adapt

```{r roads_no}
# go into mapset
ms <- "p_sam_transport_urban"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "transport_urban"

#---
# folder
gisdata_schema <- "sam_env"

#---
# public roads

# layer to use
layer_name <- "public_roads_se"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# list files
# execGRASS("v.external", input="PG:host=gisdata-db.nina.no user=bernardo.brandao dbname=gisdata", flags = "l")

# import public roads
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

#---
# private roads

# layer to use
layer_name <- "private_roads_se"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# import public roads
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")
```

### Sweden

```{r roads_sweden}
# go into mapset
ms <- "p_sam_transport_urban"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "transport_urban"

#---
# folder
gisdata_schema <- "sam_env"

#---
# public roads

# layer to use
layer_name <- "public_roads_se"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# list files
# execGRASS("v.external", input="PG:host=gisdata-db.nina.no user=bernardo.brandao dbname=gisdata", flags = "l")

# import public roads
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

#---
# private roads

# layer to use
layer_name <- "private_roads_se"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# import public roads
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")
```

## Railways

### Norway

Import vector first

```{r railways_no}
# go into mapset
ms <- "p_sam_transport_urban"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "transport_urban"

#---
# folder
gisdata_schema <- "sam_env"

#---
# railways

# layer to use
layer_name <- "railways_no"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# list files
# execGRASS("v.external", input="PG:host=gisdata-db.nina.no user=bernardo.brandao dbname=gisdata", flags = "l")

# import railways
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")
```

Then rasterize

### Sweden

```{r railways_sweden}
# go into mapset
ms <- "p_sam_transport_urban"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "transport_urban"

#---
# folder
gisdata_schema <- "sam_env"

#---
# railways

# layer to use
layer_name <- "railways_se"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# list files
# execGRASS("v.external", input="PG:host=gisdata-db.nina.no user=bernardo.brandao dbname=gisdata", flags = "l")

# import railways
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")
```

## Houses

### Norway

Import temporary files on urban and houses, rasterize them, then merge

```{r houses_urban_no}
# go into mapset
ms <- "p_sam_transport_urban"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "transport_urban"

#---
# import temporary layer with urban areas

# schema
gisdata_schema <- "sam_tmp"
# layer to use
layer_name <- "tmp_urban_no"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# list files
# execGRASS("v.external", input="PG:host=gisdata-db.nina.no user=bernardo.brandao dbname=gisdata", flags = "l")

# import railways
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

#---
# import temporary layer with houses (points)

# schema
gisdata_schema <- "sam_tmp"
# layer to use
layer_name <- "tmp_houses_not_urban_no"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# list files
# execGRASS("v.external", input="PG:host=gisdata-db.nina.no user=bernardo.brandao dbname=gisdata", flags = "l")

# import railways
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

#-----------------
# rasterize

# rasterize urban areas with value 10
rgrass7::execGRASS("g.region", raster = "master_grid_100m_norway@p_sam_tools", 
                   flags = "p")
rgrass7::execGRASS("v.to.rast", input = "tmp_urban_no", type = "area",
                   output = "tmp_urban_x10_no", use = "val", value = 10)

# rasterize houses counting the number of features per pixel
rgrass7::execGRASS("g.region", raster = "master_grid_100m_norway@p_sam_tools", 
                   flags = "p")
# rgrass7::execGRASS("r.mask", flags = "r")
oneimpact::grass_v2rast_count(x = "tmp_houses_not_urban_no", 
                              output = "tmp_houses_not_urban_count_no", 
                              set_region = FALSE)

#-----------------
# patch results

# region
rgrass7::execGRASS("g.region", raster = "master_grid_100m_norway@p_sam_tools", 
                   flags = "p")
# mask
rgrass7::execGRASS("r.mask", vector = "norway_limits@p_sam_tools")
# patch
rgrass7::execGRASS("r.patch", input = "tmp_urban_x10_no,tmp_houses_not_urban_count_no",
                   output = "houses_count_no")
# export for checking
dir <- "../data/sam_env/transport_urban/houses/processed/"
rgrass7::execGRASS("r.out.gdal", input = "houses_count_no", 
                   output = paste0(dir, "houses_count_no.tif"),
                   createopt = "COMPRESS=DEFLATE")
# remove mask
rgrass7::execGRASS("r.mask", flags = "r")

```

Import vector file, then rasterize

```{r houses_sweden}
# go into mapset
ms <- "p_sam_transport_urban"
execGRASS("g.mapset", mapset = ms)

#---
# settings
type_of_info <- "transport_urban"

#---
# folder
gisdata_schema <- "sam_env"

#---
# houses

# layer to use
layer_name <- "houses_se"
# schema and layer
sch_lay <- paste(gisdata_schema, layer_name, sep = ".")

# list files
# execGRASS("v.external", input="PG:host=gisdata-db.nina.no user=bernardo.brandao dbname=gisdata", flags = "l")

# import railways
execGRASS("v.in.ogr", input = str_con, layer = sch_lay, 
          output = layer_name, flags = "overwrite")

#--------------------
# rasterize houses
# rasterize houses counting the number of features per pixel
rgrass7::execGRASS("g.region", vector = "herding_districts_se_union@p_sam_tools",
                   align = "master_grid_100m_norway@p_sam_tools", 
                   flags = "p")
# mask
rgrass7::execGRASS("r.mask", vector = "herding_districts_se_union@p_sam_tools")
# rgrass7::execGRASS("r.mask", flags = "r")
oneimpact::grass_v2rast_count(x = "houses_se", 
                              output = "houses_count_se", 
                              set_region = FALSE, 
                              overwrite = TRUE)
# export for checking
dir <- "../data/sam_env/transport_urban/houses/processed/"
rgrass7::execGRASS("r.out.gdal", input = "houses_count_se", 
                   output = paste0(dir, "houses_count_se.tif"),
                   createopt = "COMPRESS=DEFLATE")
# remove mask
rgrass7::execGRASS("r.mask", flags = "r")
```

We must now patch the maps for Norway and Sweden

Virtual raster?




